---
title: "Exploratory Data Analysis"
subtitle: "NFL"  
author: 
  - "Sarah Sult"
  - "Brooke Coneeny"
  - "Devin Basley"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  fig.show = TRUE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#1381B0",
  secondary_color = "#1381B0",
  inverse_header_color = "#FFFFFF"
)
```


class: inverse center middle

# Background


---

# Data

*   From nflfastR package
*   2020 regular season passing plays
--

*   Each observation is one pass from this season
--

*   Variables
--

    *   Information on the player attempting the pass
--

    *   Info about the pass
--

    *   Context about the reciever
--

    *   Context about the team
--

    *   Game context
--

    *   Description about the play

---

<span style="color: white;">
Filling up space </span>

<span style="color: white;">
Filling up space </span>

<span style="color: white;">
Filling up space </span>


```{r, echo=FALSE, fig.align='center'}
library(gt)
nfl_passing_plays %>%
  select(passer_player_name, posteam, qb_hit,sack, interception, complete_pass, yards_gained, epa) %>%
  slice(373:378) %>%
  gt()
```


---

##Comparing NFL Team's Offensive Line Strength

```{r plot1, fig.align = "center", echo=FALSE}
nfl_passing_plays %>%
  group_by(posteam) %>%
  summarize(total_hit = sum(qb_hit) + sum(sack)) %>%
  ungroup() %>%
  mutate(posteam = fct_reorder(posteam, total_hit)) %>%
  ggplot(aes(x = posteam, y = total_hit)) + 
  geom_col(fill = "slategray3") +
  labs(
    x = "Possessing Team",
    y = "QB Hits + Sacks",
    caption = "Data courtesy of nflfastR"
  ) +
  theme_bw() +
  theme(
    plot.background = element_rect(fill = "grey95"),
    panel.background = element_rect(fill = "grey95"),
    panel.grid.minor = element_blank(),
    text = element_text(family = "Century", size = 12),
    plot.title = element_text(size = 17),
    axis.text.x = element_text(size = 8, angle = 90)
  )
```


---

#Interception Not More Likely When Hit

```{r,echo=FALSE}
nfl_passing_plays %>%  
  filter(sack != 1) %>%                 #remove sacks because it artificially inflates 
                                      #the value of hit/no interception
  group_by(qb_hit, interception, complete_pass) %>%
  summarize(count = n(),
            joint_prob = count / nrow(nfl_passing_plays)) %>%
  ungroup() %>%
  mutate(qb_hit_name = case_when(
    qb_hit == 0 ~ "No hit",
    TRUE ~ "Hit"
  ), interception_name = case_when(
    interception == 0 ~ "No interception",
    TRUE ~ "Interception"
  ), complete_pass = case_when(
    complete_pass == 0 ~ "Incomplete pass",
    TRUE ~ "Complete pass"
  )) %>%
  ggplot(aes(x=qb_hit_name, y=interception_name)) +
  geom_tile(aes(fill=count), color="white")+
  geom_text(aes(label = round(joint_prob, digits=4)), color = "white")+
  facet_wrap(~ complete_pass, ncol=2) +
  scale_fill_viridis_b()+
  theme_bw()+
  theme(legend.position = "bottom", legend.key.width = unit(2,"cm")) +
  labs(
    x = "QB Hit",
    y = "Interception",
    caption = "Data courtesy of nflfastR"
  )
```


---

#Passing Plays Don't Get Many Yards

```{r, echo = FALSE}
((yards_gained_dens_total + theme(axis.title.x = element_blank())+ yards_gained_dens + theme(axis.title.y = element_blank(), axis.title.x = element_blank(), legend.position = "none")) /( yards_gained_ecdf_total + yards_gained_ecdf + theme(axis.title.y = element_blank()))) + 
  plot_layout(guides="collect") + plot_annotation(caption = "Data courtesy of nflfastR")

```


---
#Methods for Clustering Example
*   Collapsed the data into one row per thrower with their total hits (hits + sacks), total expected points added, and total number of throws
--

*   Graphed the ecdf of total throws to determine our cutoff
--

```{r, echo=FALSE}
nfl_passing_plays_total_hits_and_epa %>%
  ggplot(aes(x=total_throws)) +
  stat_ecdf()+
  geom_vline(xintercept = 5, linetype = "dashed", color = "darkred")+
  theme_bw()+
  labs(x = "Total Throws",
       y = "Proportion")
```

---

##Categorizing Players by Total Hits and Total EPA

```{r, echo = FALSE}
nfl_passing_plays_total_hits_and_epa_over5 %>%
  mutate(player_clusters = as.factor(cutree(hits_epa_complete_hclust, k=3)),) %>%
  ggplot(aes(x=total_hits, y=total_epa, color = player_clusters))+
  geom_point(alpha = .75)+
  theme_bw()+
  theme(legend.position = "bottom")+
  ggthemes::scale_color_colorblind()+
  labs(x = "Total Hits",
       y = "Total Expected Points Added",
       caption = "Data courtesy of nflfastR")

```

